---
title: "AfricaCountryBins"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.path = "man/figures/README-",
  collapse = TRUE
)
```

  <!-- badges: start -->
  [![R-CMD-check](https://github.com/delabj/AfricaCountryBins/workflows/R-CMD-check/badge.svg)](https://github.com/delabj/AfricaCountryBins/actions)

Create Uniform Square Country Cartogram heatmaps

## Description

Heatmaps in a tile shape are an alternative to choropleth maps that don't distort based on land mass size.

This package is a essentially a fork of [Bob Rudis's statebins package](https://github.com/hrbrmstr/statebin) designed for use for data involving Africa. 

This grid attempts to preserve the general position of countries on the contienet, while providing an equal area by country.

## Whatcha Get?

The main function implemented is 
 - `geom_countrybins_africa`: which is a countrybin Geom for Africa
 
Sample dataset `africa_mines` (see `?africa_mines` for details)

## Installation

Use the following to install: 
```{r eval=FALSE}
devtools::install_github("delabj/AfricaCountryBins")
# or 
remotes::install_github("delabj/AfricaCountryBins")
```

To use these options you'll have to install devtools or remotes respectively. 

## Examples

```{r ex-01, fig.width=600/96, fig.height=600/96}
library(AfricaCountryBins)
library(hrbrthemes)
library(tidyverse)

data("africa_mines")

count(africa_mines, countrycode, wt=count) %>%
  mutate(n = ifelse(n == 0, NA, n)) %>%
  ggplot(aes(country = countrycode, fill = n)) +
  geom_countrybins_africa() +
  coord_fixed() +
  scale_fill_viridis_c(
    name = "# mines (log2)",
    option = "magma",
    direction = -1,
    na.value = "#DEE5E8",
    trans = "log2",
    label = scales::comma_format(1)
  ) +
  guides(
    fill = guide_colourbar(
      title.position = "top"
    )
  ) +
  labs(
    title = "Mines in Africa",
    caption = "Source: data.world <hdata.world/datainspace/mines-in-africa>"
  ) +
  theme_ipsum_es(grid="") +
  theme(legend.position = c(0.2, 0.4)) +
  theme(legend.direction = "horizontal") +
  theme(legend.key.width = unit(1.5, "lines")) +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r ex-02, fig.width=1200/96, fig.height=1200/96}
africa_mines %>%
  count(countrycode, mine_type, wt = count) %>%
  mutate(mine_type = fct_reorder(mine_type, n, sum)) %>%
  filter(mine_type %in% rev(levels(mine_type))[1:8]) %>%
  mutate(n = ifelse(n == 0, NA, n)) %>%
  ggplot(aes(country = countrycode, fill = n)) +
  geom_countrybins_africa(
    radius = unit(0, "pt"),
    family = font_an, size = 2
  ) +
  scale_fill_viridis_c(
    name = "# mines (log2)",
    option = "magma",
    direction = -1,
    na.value = "#DEE5E8",
    trans = "log2",
    label = scales::comma_format(1)
  ) +
  coord_equal() +
  facet_wrap(~mine_type) +
  guides(
    fill = guide_colourbar(
      title.position = "top"
    )
  ) +
  labs(
    title = "Mines in Africa",
    caption = "Source: data.world <hdata.world/datainspace/mines-in-africa>"
  ) +
  theme_ipsum(grid="", strip_text_face = "bold") +
  theme(legend.position = c(0.85, 0.2)) +
  theme(legend.direction = "vertical") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank())
```
